Covariance as constraint: Leveraging leaf economic relationships to constrain plant functional type trait estimates

### 2017-11-16 manuscript-md/02_methods.Rmd
Finished manuscript draft

Foliar trait data for this analysis comes from the TRY global traits database
(@kattge_try_2011; see Supplementary Table 1 for a complete table of component datasets and references.).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, mm$^2$ mg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at leaf temperature ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
We performed analyses separately for both mass- and area-normalized traits to avoid artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification (see @kattge_try_2011).

@@

Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.

@@

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model (CLM4.5, @clm45_note; Tab. \@ref(tab:pfts), Fig. \@ref(fig:samplesize)).
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1 km^{2}$ ) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean temperature for each species,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating missing values in the data ($x$).
Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows: For a block of data $\mathbf{x'}$ containing missing observations in the columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$...

@@

For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].

We implemented this sampling algorithm in R [version 3.4.2; @rstats].
The code for fitting these models to arbitrary data is available as an open source R package (`mvtraits`) at <http://github.com/ashiklom/mvtraits>.

@@

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.~\@ref(fig:meancompare)).
In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.~2 and 3).

@@

If all three models perform equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
On the other hand, models that perform better should have lower slope ($b_1$) coefficients, indicating a slower increase in the uncertainty ($\alpha$) with decreasing sample size ($n$).

To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.~\@ref(fig:stickpairs)).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.~\@ref(fig:corrbars)).

>>>>

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, mm$^2$ mg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at leaf temperature ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
We performed analyses separately for both mass- and area-normalized traits to avoid potential artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

@@

Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @onoda_2011_global; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.

```{r "trydata"}
crtable <- readRDS("extdata/crossref_table.rds") %>%
  mutate(
    Reference = case_when(
      Reference == "@try_data_03" ~ "@baraloto_decoupled_2010",
      Reference == "@try_data_09" ~ "@wright_worldwide_2004",
      Reference == "@try_data_14" ~ "@kattge_2009_vcmax",
      Reference == "@try_data_15" ~ "@laughlin_multi-trait_2010",
      Reference == "@try_data_26" ~ "@onoda_2011_global",
      Reference == "@try_data_30" ~ "@messier_how_2010",
      Reference == "@try_data_33" ~ "@pierce_allocating_2013",
      Reference == "@try_data_41" ~ "@poorter_causes_2009",
      Reference == "@try_data_44" ~ "@choat_global_2012",
      Reference == "@try_data_59" ~ "@laughlin_climatic_2011",
      Reference == "@try_data_67" ~ "@wright_does_2012",
      Reference == "@try_data_80" ~ "@choat_global_2012",
      Reference == "@try_data_114" ~ "@kichenin_2013_contrasting",
      Reference == "@try_data_134" ~ "@diaz_plant_2004",
      TRUE ~ Reference
    )
  )
cap <- "Data sources and observation counts for foliar trait data"
kable(crtable, caption = cap)
```

@@

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ \@ref(fig:samplesize)].
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species or more likely to occur in nature.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1 km^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

@@

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x'}[m]$ are drawn randomly from a marginal multivariate normal distribution at each iteration of the sampling algorithm...

@@

For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
We implemented this sampling algorithm in an open source, publicly available R [version 3.4.2; @rstats] package (<http://github.com/ashiklom/mvtraits>).

@@

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.\ \@ref(fig:meancompare)).
In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.\ 3 and 4).

@@

If all three models performed equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
On the other hand, models that perform better should have lower slope ($b_1$) coefficients, indicating a slower increase in the uncertainty ($\alpha$) with decreasing sample size ($n$).

To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ \@ref(fig:corrbars)).

### 2017-11-16 manuscript-md/03_results.Rmd
Finished manuscript draft

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig. \@ref(fig:meancompare)).
Where we observed differences between modles, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

@@

C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.

C3 grasses also had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had among the highest $J_{max,area}$ but the lowest $J_{max,mass}$.

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
For some of these trait-PFT combinations, such as estimates $J_{max,mass}$ for shrub PFTs, the additional constraint from trait covariance substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
However, this analysis revealed no consistent significant benefit from the hierarchical model for our specific analysis.

@@

Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig. \@ref(fig:stickpairs)).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
On the other hand, correlations between leaf lifespan and area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive both globally and for evergreen shrubs, but was negative for broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.

A large number of pairwise trait correlations were not significant.
In some cases, this was driven by sample size (Fig. \@ref(fig:samplesize)).
For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

@@

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig. \@ref(fig:corrbars)).
The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
Correlation strength was often related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

>>>>

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ \@ref(fig:meancompare)).
Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

@@

C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.

C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
However, this analysis revealed no consistent significant benefit from the hierarchical model.

@@

Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs but negative for broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.
The correlation between leaf lifespan and $R_{d,area}$ significant and positive globally, but was not significant within any PFTs.
The correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees, but not significant for any other PFTs or globally.

A large number of pairwise trait correlations were not significant.
In some cases, this was driven by sample size (Fig.\ \@ref(fig:samplesize)).
For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

@@

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).
The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

### 2017-11-16 manuscript-md/index.Rmd
Finished manuscript draft

---
title: "Covariance as constraint: Leveraging leaf economic relationships to constrain plant functional type trait estimates"
author: >
  Alexey N. Shiklomanov$^1$,
  Elizabeth M. Cowdery$^1$,
  Michael Bahn$^2$,
  Sabina Burrascano$^3$,
  Chaeho Byun$^4$,
  Joseph Craine$^5$,
  Andrés Gonzalez-Melo$^6$,
  Alvaro G. Gutiérrez$^7$,
  Steven Jansen$^8$,
  Nathan Kraft$^9$,
  Koen Kramer$^{10}$,
  Vanessa Minden$^{11,12}$
  Ülo Niinemets$^{13}$,
  Yusuke Onoda$^{14}$,
  Enio Egon Sosinski$^{15}$,
  Nadejda A. Soudzilovskaia$^{16}$,
  Michael C. Dietze$^1$
bibliography:
  - library.bib
  - extra.bib
csl: new-phytologist.csl
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
---
$^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
$^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
$^3$Department of Environmental Biology, Sapienza University of Rome, Piazzale Aldo Moro 5, 00185, Rome, Italy;
$^4$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
$^5$Jonah Ventures, Manhattan, KS 66502, USA;
$^6$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
$^7$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
$^8$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
$^9$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
$^{10}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
$^{11}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
$^{12}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
$^{13}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
$^{14}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
$^{15}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
$^{16}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands

@@

* We investigated whether global leaf economic relationships across plant functional types (PFTs) can be reconciled with within-PFT relationships, and the extent to which this hierarchical structure can be used to constrain trait estimates.

@@

* Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
Sometimes, the multivariate model significantly shifted trait means, resulting in estimates better aligned with patterns predicted from global ecology and biogeography.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

>>>>

---
title: "Covariance as constraint: Leveraging leaf economic relationships to constrain plant functional type trait estimates"
author: >
  Alexey N. Shiklomanov$^1$,
  Elizabeth M. Cowdery$^1$,
  Michael Bahn$^2$,
  Sabina Burrascano$^3$,
  Chaeho Byun$^4$,
  Joseph Craine$^5$,
  Andrés Gonzalez-Melo$^6$,
  Alvaro G. Gutiérrez$^7$,
  Steven Jansen$^8$,
  Nathan Kraft$^9$,
  Koen Kramer$^{10}$,
  Vanessa Minden$^{11,12}$
  Ülo Niinemets$^{13}$,
  Yusuke Onoda$^{14}$,
  Enio Egon Sosinski$^{15}$,
  Nadejda A. Soudzilovskaia$^{16}$,
  Michael C. Dietze$^1$
bibliography:
  - library.bib
  - extra.bib
  - tryrefs.bib
csl: new-phytologist.csl
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
---
$^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
$^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
$^3$Department of Environmental Biology, Sapienza University of Rome, Piazzale Aldo Moro 5, 00185, Rome, Italy;
$^4$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
$^5$Jonah Ventures, Manhattan, KS 66502, USA;
$^6$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
$^7$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
$^8$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
$^9$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
$^{10}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
$^{11}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
$^{12}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
$^{13}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
$^{14}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
$^{15}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
$^{16}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands

@@

* We investigated whether global leaf economic relationships across plant functional types (PFTs) area also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.

@@

* Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.
